\textbf{COMP0086 Summative Assignment}

### 2022-11-11 submission.tex
comment code and refactor

             We can calculate:
             $$f'(x) = \frac{\partial f(x, a)}{\partial x}= \exp(x)$$
             Assuming we can evaluate $\exp(x)$, our update equation:
             $$x_{n+1} = x_{n} - \frac{\exp(x_n) - a}{\exp(x_n)}$$
             Simplifying:
             $$x_{n+1} = x_{n} + \frac{a}{\exp(x_n)} - 1$$
\end{enumerate}

>>>>

             We can calculate:
             $$f'(x) = \frac{\partial f(x, a)}{\partial x}= \exp(x)$$
             Assuming we can evaluate $\exp(x)$, our update equation:
             $$x_{n+1} = x_{n} - \frac{\exp(x_n) - a}{\exp(x_n)}$$
             Simplifying:
             $$x_{n+1} = x_{n} + \frac{a}{\exp(x_n)} - 1$$
\end{enumerate}
\end{enumerate}

\newpage
\section*{Question 8}
\begin{enumerate}
\item[(a)] For:
$$\sup_{\{\textbf{x}\in \mathbb{R}^n\}} R_A(\textbf{x})$$
where $R_A(\textbf{x}) = \frac{\textbf{x}^T\textbf{Ax}}{\|\textbf{x}\|^2}$, we want to show that a maximum is attained.

To do this, we will show the above optimisation can be equivalently formulated:
$$\sup_{\{\textbf{x}\in \mathbb{R}^n | \| \textbf{x}\|=1\}} R_A(\textbf{x})$$
We begin by considering any $\textbf{w} \in \mathbb{R}^n$ and let $\textbf{x} = \frac{\textbf{w}}{\|\textbf{w}\|}$. Because $\|\textbf{x}\| = 1$ we can substitute:
$$\sup_{\{\frac{\textbf{w}}{\|\textbf{w}\|}\in \mathbb{R}^n | \| \frac{\textbf{w}}{\|\textbf{w}\|}\|=1\}} R_A(\textbf{x}) = \sup_{\textbf{x}\in \mathbb{R}^n | \| \textbf{x}\|=1} \frac{\textbf{w}^T\textbf{Aw} \|\textbf{w}\|^2}{\|\textbf{w}\|^2 \textbf{w}^T\textbf{w}}$$
where $\|\textbf{x}\|^2 = \textbf{x}^T\textbf{x}$.

The set $\{\frac{\textbf{w}}{\|\textbf{w}\|}\in \mathbb{R}^n | \| \frac{\textbf{w}}{\|\textbf{w}\|}\|=1\}$ holds for all $\textbf{w} \in \mathbb{R}^n$ so we can rewrite:
$$\sup_{\{\textbf{w}\in \mathbb{R}^n\}} \frac{\textbf{w}^T\textbf{Aw} \|\textbf{w}\|^2}{\|\textbf{w}\|^2 \textbf{w}^T\textbf{w}}$$
We can simplify the expression:
$$\sup_{\{\textbf{w}\in \mathbb{R}^n\}} \frac{\textbf{w}^T\textbf{Aw} }{\textbf{w}^T\textbf{w}}$$
$$\sup_{\{\textbf{w}\in \mathbb{R}^n\}} R_A(\textbf{w})$$

and recover our original optimisation problem by letting $\textbf{x}=\textbf{w}$, showing that it is equivalent to the supremum over the unit sphere. Assuming the set containing the unit sphere is compact, the extreme value theory of calculus states that $\sup_{\{\textbf{x}\in \mathbb{R}^n | \| \textbf{x}\|=1\}} R_A(\textbf{x})$ is attained so equivalently $\sup_{\{\textbf{x}\in \mathbb{R}^n\}} R_A(\textbf{x})$ is attained as required.

\item[(b)] We can now reformulate the optimisation as:
$$\sup_{\{\textbf{x}\in \mathbb{R}^n | \| \textbf{x}\|=1\}} \frac{\textbf{x}^T\textbf{Ax}}{\|\textbf{x}\|^2}$$
But because $\| \textbf{x}\|=1$, we can equivalently write:
$$\sup_{\{\textbf{x}\in \mathbb{R}^n | \| \textbf{x}\|=1\}} \textbf{x}^T\textbf{Ax}}$$
Thus, showing $\textbf{x}^T\textbf{Ax}} \leq \lambda_1$ will be equivalent to showing $R_A(\textbf{x}) \leq \lambda_1$ for $\|\textbf{x}\|=1$.
We know that for all $\textbf{x} \in \mathbb{R}^n$:
$$\textbf{x} = \sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i$$
so we can write:
$$\textbf{x}^T\textbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\textbf{A}\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i\bigg)$$
Given that $\xi_i$ are eigenvectors of $\textbf{A}$ corresponding to eigenvalues $\lambda_i$:
$$\textbf{x}^T\textbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\bigg(\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x}) \xi_i\bigg)$$
Given that the eigenvectors $\xi_i$ form an orthonormal basis, we know that $\xi_i^T\xi_j = 0$ when $i\neq j$ and $\xi_i^T\xi_j = 1$ when $i=j$, so:
$$\textbf{x}^T\textbf{Ax} =\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$$
We know that $\|\textbf{x}\|^2 = 1$ so $\|\textbf{x}\|^2 = \sum_{j=1}^{n} x_i^2 = \sum_{j=1}^{n} (\xi_j \textbf{x})^2 = 1$. Thus the quantity $\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$ is simply a weighted average of $\lambda_i$'s and we can write:
$$\textbf{x}^T\textbf{Ax} =\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2 \leq \lambda_1$$
where $\lambda_1$ is the largest eigenvalue of eigenvalues $\lambda_i$. Therefore, $R_A(\textbf{x}) \leq \lambda_1$ as required.

\item[(c)] Given that $\lambda_j < \lambda_1 \forall j > k$ and $\textbf{x} \in span\{\xi_{k+1}, ..., xi_{n}\}$, we can rewrite $\textbf{x}$:
$$\textbf{x} = \sum_{i=k+1}^n (\xi_i^T \textbf{x}) \xi_i$$
From the same argument as (b) we can bound $\textbf{x}^T\textbf{Ax}}$:
$$\textbf{x}^T\textbf{Ax}} = \sum_{i=k+1}^n \lambda_i(\xi_i^T \textbf{x})^2 \leq \max\{\lambda_{k+1}, ..., \lambda_n\}$$
But given that the maximum eigenvalue $\lambda_1$ is not contained in $\{\lambda_{k+1}, ..., \lambda_n\}$:
$$\max\{\lambda_{k+1}, ..., \lambda_n\} < \lambda_1$$
and therefore,
$$R_A(\textbf{x}) < \lambda_1$$
as required.
